|

### 2019-08-19 article/article.Rmd
mignozzetti-commit

---
title:  |
        | Bottom-Up Accountability and Public Service Provision: Evidence from a Field Experiment in Brazil
author: 
- "Danilo Freire[^freire]"
- "Manoel Galdino[^galdino]"
- "Umberto Mignozzetti[^mignozzetti]"
date: \today
thanks: "We thank Fábio Barros, Guilherme Jardim Duarte, Stephen Herzog, Catarina Roman and David Skarbek for their helpful comments. We are specially grateful to the staff of Transparência Brasil for their excellent research assistance. This research was pre-registered on the EGAP pre-registry tool (<https://egap.org/registration/4505>) and approved by the Brazilian Ministry of Education. The authors thank the 2016 Google Social Impact Challenge for financial support and declare there are no conflict of interest. Replication data and code are available at <http://github.com/umbertomignozzetti/tdp-accountability>."
abstract: "Does grassroots participation improve public service delivery? We study the effect of a mobile phone application that allows citizens to monitor school construction projects in 1030 Brazilian municipalities. The app provides a platform where users can submit photos of construction sites, consult independent engineers, and contact their representatives about project delays. Our results show that the app has a null impact on public spending and school completion times. Additionally, we find that politicians do not react to individual requests. The results question the impact of local monitoring on public service performance and suggest that interventions targeted at other groups may produce better policy outcomes."
abstractspacing: double
keywords: accountability, Brazil, impact evaluation, state capacity, technology
fontsize: 12pt
margin: 
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
---

@@

One of the key determinants of efficient public services provision is a robust accountability system [@besley2003incentives; @cameron2004public; @ferejohn1986incumbent; @o1998horizontal]. In its standard definition, accountability is understood as the process of holding authorities responsible for their actions [@finer1941administrative; @mulgan2000accountability; @o1990bureaucratic]. Past studies show that accountability has a positive impact on governance, as it ensures that politicians act on behalf of voters [@freire2010ngp; @moncrieffe1998reconceptualizing], reduces the opportunities for rent-seeking and corruption [@deininger2005does; @wenar2006accountability] and improves the quality of public services [@adsera2003you; @bjorkman2009power]. Recent research also suggests that accountability leads to higher economic growth because it limits state discretion in the economy and increases long-term investments in human capital [@benhabib2010economic; @suebvises2018social; @ponzetto2018social].

However, accountability systems take many forms. One promising model is that of _bottom-up monitoring_, in which citizens receive information about the shortcomings of a given project so they can evaluate and pressure underperforming public officials [@kosack2014does; @molina2016community; @raffler2018weakness]. Proponents argue bottom-up accountability is effective because: 1) constituents have first-hand information about the outcomes of local policies; 2) citizens have incentives to attack corruption that directly affects them; 3) policy-makers are sensitive to social punishment from their own communities [@serra2011combining, 570]. In this regard, bottom-up accountability offers a potential solution to the principal-agent dilemma in public service by aligning the interests of state officials with those of the constituency they serve [@barro1973control; @raffler2018weakness, 2].

Here we assess the impact of _Tá de Pé_ (TDP), a mobile phone application designed to lower the costs of evaluating public works and punishing political representatives in Brazil. Developed by Transparência Brasil[^transparencia], TDP allows Brazilian citizens to know the location of public school construction sites, check their completion status and anonymously request information from competent authorities. TDP users can also take pictures of the construction sites and submit them to independent engineers for examination. If the engineers classify the construction as delayed, TDP prompts users to send a message to the mayor's office asking for completion estimates and explanations about the construction status. TDP has been online since April 2017 in 1030 Brazilian municipalities and was the winner of the 2016 Google Social Impact grant with more than 200,000 popular votes[^google].

@@

We use the TDP app to conduct two experimental interventions and test its impact on six outcomes related to school completion rates and complaints to public authorities. Overall, our results show that providing information to citizens has no consistent impact on policy outcomes. We find that the TDP app increased the likelihood of school completion by 7 percent in the first experiment, but the effect is null in the second intervention. We also find that the app increased the percentage of the executed budget by 157.5 percent in our second experiment, but only after nearest neighbourhood matching adjustment. Importantly, both interventions have no significant impact on any of the other measures of school investment and completion rates we employed in this study.

The results raise questions about the ability of citizens to hold representatives accountable using bottom-up strategies. Although studies like @bjorkman2009power, @fiala2017social and @reinikka2005fighting report better policy outcomes after providing information to local communities, our findings are closer to @banerjee2010pitfalls, @keefer2014mass, @lieberman2014does, @bjorkman2017experimental, @olken2007monitoring and @raffler2018weakness and offer little evidence that information interventions lead to greater government responsiveness. Our analysis suggests that grassroots monitoring does not affect government behaviour in Brazil, and other models of political accountability may be more effective for public service provision.

@@

The _Tá de Pé_[^tdp] (TDP) cell phone application is an initiative carried out by Transparência Brasil to foster bottom-up accountability in the Brazilian public sector. More specifically, the main goal of the TDP project is to improve responsiveness in government education expenditures. The TDP app incentivises citizens to provide up-to-date information about unfinished schools in their neighbourhoods, and that information will be assessed by a group of independent specialists. In the case the construction is behind schedule, TDP provides a writing platform whereby citizens can contact their representatives quickly and anonymously. The app then prompts citizens to write a notification to the mayor's office, which has 15 days to reply. If they do not respond to the request, the app forwards the notification to the Brazilian federal government, making it harder for the municipality to access federal funds in the future. The motivation for the intervention is that information delivery causes individuals to closely monitor public works. This, in turn, results in better social outcomes as public agents become more responsive to community demands.

@@

We built the app in the first three months of 2017 and tested it in May 2017. The first stable Android version was deployed on Google Play on 14 August 2017. A version for iOS came about six months later. In October 2017, we started a Facebook campaign in order to publicise the app. Facebook is the most widely used social media network in Brazil with around 72 million users in 2017 [@statista2019facebook]. The campaigns cost US$ 4,479.29 and Facebook advertisements attracted 2,028 new users to the platform in October 2017 only.

@@

\begin{figure}
     \begin{minipage}[t]{.3\textwidth}
       \centering
       \includegraphics[scale=0.25]{tdp03.jpg}
       \label{}
     \end{minipage}
     \hfill
     \begin{minipage}[t]{.3\textwidth}
       \centering
       \includegraphics[scale=0.25]{tdp01.jpg}
     \end{minipage}
     \hfill
     \begin{minipage}[t]{.3\textwidth}
       \centering
       \includegraphics[scale=0.25]{tdp02.jpg}
     \end{minipage}
\caption{The \textit{Tá de Pé} mobile phone application. The first image presents a list of the closest school construction sites. The second image shows that a school construction is delayed by 9 months. The last image shows how citizens can add information to the photos they submit via the app.} 
\label{fig:screenshots}
\end{figure}

Transparência Brasil partnered with the Brazilian branch of _Engineers without Borders_ (EWB), an independent non-governmental organisation[^ewb], to provide technical assessment of school completion rates based on user-submitted photos and GPS coordinates. The engineers' reports are later uploaded to the TDP database and stored on the users' mobile devices so citizens can follow the progress of the constructions.

@@

The TDP has also received feedback from Brazilian computer programmers and policy analists. In 2017 and 2018, Transparência Brasil announced two team programming competitions, called 'Tá de Pé Hackathons', where contributors could fix code bugs and suggest new functionalities to the TDP project. One of these innovations consists of a Twitter bot (<https://twitter.com/tadepeapp>) which posts a message on the social network each time a user submits a new picture for evaluation. This allows any interested parties, including those who do not use the TDP app, to check the state of school construction sites.

@@

Between August 2017 and August 2018, we implemented two interventions to measure the effect of the TDP app on six school construction outcomes. The outcomes are: 1) a placebo outcome indicating the percentage of the investment executed before the impact evaluation started; 2) the percentage of public investment allocated to school constructions; 3) the difference between the percentage invested before and after the interventions; 4) the number of finished constructions; 5) the number of cancelled constructions; 6) the number of updated conclusion dates. We expect the first outcome to remain unchanged and the five remaining ones to increase.

The first intervention was carried out from August 2017 to February 2018 using the Android version of TDP. The unit of analysis for this experiment is the Brazilian municipality. We randomly selected `r as.numeric(table(impEvalph1$treatOriginal)[1])` municipalities to the control group and included `r as.numeric(table(impEvalph1$treatOriginal)[2])` in the treatment group. Our control condition consists in removing all information about school construction from the TDP app in the chosen municipalities, so that citizens had no incentive to use the app.

@@

Table \ref{tab:results1} summarises the main results of intervention 1. Each column represents the treatment effect of the TDP app on one of the six outcomes we measured for this study. All models reported here include the five control variables described in the previous section and Brazilian states' fixed effects. We also estimated the models without control variables, without fixed effects and with nearest-neighbour matching. The results are very similar to those below.

@@

We find that the intervention is sufficient to improve school completion rates. The TDP application increases in `r round(100*resInt1[[2]][[16]]$coefficients[1], 1)` percent the chance that a school construction is reported as finished by the end of the intervention. This is in line with our theoretical expectations, but as we discuss below, the finding is not consistent. All other coefficients are not statistically significant at conventional levels. On the one hand, the results indicate that our placebo outcome, the percentage of the invested executed before the intervention, indeed behaves as predicted. On the other hand, we expected the four remaining outcomes to increase after the introduction of the app. The literature on bottom-up accountability argues that more information about the shortcomings of public service provision will mobilise citizens who, in turn, would pressure state agents to produce better social policies. Our results do not lend support to that hypothesis.

We observe a similar pattern in the second intervention. Table \ref{tab:results2} shows that the treatment does not have a statistically significant effect on any our outcomes of interest, including the placebo. This provides evidence that our previous findings are not a statistical fluke, and questions the robustness of the effect of the TDP app on school completion. Nevertheless, in our additional tests we find that the app improved the difference between the baseline and the end-line reported investment by 157.5 percent. This is a large effect, yet it only reaches statistical significance when we employ inverse probability weights with block randomisation and one-to-one nearest neighbourhood matching.

@@

We also note that the effect sizes are small in each of our original models. Therefore, even if the treatment effects were statistically significant, the TDP app would have a reduced impact over the outcomes. This has an important substantive meaning, as it poses a direct challenge to the bottom-up accountability theory, at least for the school construction outcomes we estimated here.

@@

What factors, then, are driving these results? It seems unlikely that the null results derive from flaws in the research design. First, our study is well powered. We included 1000 municipalities in the treatment groups, thus we should be able to capture small average treatment effects if they exist. Second, balance and manipulation tests indicate that the treatment allocation was successful, so we can rule out problems in the randomisation procedures. Third, after doing a series of robustness tests --- including matching and randomised inference --- we still find no firm evidence of treatment effect.

@@

In sum, our experiments suggest that popular participation and bottom-up monitoring may not be effective to improve public service delivery. Nevertheless, the null findings are informative to researchers and policy-makers. The most important recommendation derived from this study is that interventions targeting elite groups, such as lobbyists or civil servants, may render better policy outcomes than those focused at the grassroots levels. Another core lesson is that although digital interventions are promising means to deliver information, perhaps they do not have the same impact as personal, face-to-face communication. Since many developing countries share Brazil's issues with education provision, the shortcomings we describe here serve as warnings for future interventions.

>>>>

---
title:  |
        | Bottom-Up Accountability and Public Service Provision: Evidence from a Field Experiment in Brazil
author: 
- "Danilo Freire[^freire]"
- "Manoel Galdino[^galdino]"
- "Umberto Mignozzetti[^mignozzetti]"
date: \today
thanks: "We thank Fábio Barros, Guilherme Jardim Duarte, Stephen Herzog, Catarina Roman and David Skarbek for their helpful comments. We are specially grateful to the staff of Transparência Brasil for their excellent research assistance. This research was pre-registered on the EGAP pre-registry tool (<https://egap.org/registration/4505>) and approved by the Brazilian Ministry of Education. The authors thank the 2016 Google Social Impact Challenge for financial support and declare there are no conflict of interest. Replication data and code are available at <http://github.com/umbertomignozzetti/tdp-accountability>."
abstract: "Does grassroots participation improve public service delivery? We study the effect of a mobile phone application that allows citizens to monitor school construction projects in 1030 Brazilian municipalities. The app provides a platform where users can submit photos of construction sites, consult independent engineers, and contact mayor's office about project delays. Our results show that the app has a null impact on the school construction outcomes. Additionally, we find that politicians are unresponsive to individual requests. The results question the impact of local monitoring on public service performance and suggest that interventions targeted at other groups may produce better policy outcomes."
abstractspacing: double
keywords: accountability, Brazil, impact evaluation, state capacity, technology
fontsize: 12pt
margin: 
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
---

@@

One of the key determinants of efficient public services provision is a robust accountability system [@besley2003incentives; @cameron2004public; @ferejohn1986incumbent; @o1998horizontal]. [XXXX A robust accountability system is crucial for efficient public services provision XXXX] In its standard definition, accountability is understood as the process of holding authorities responsible for their actions [@finer1941administrative; @mulgan2000accountability; @o1990bureaucratic]. Past studies show that accountability has a positive impact on governance, as it ensures that politicians act on behalf of voters [@freire2010ngp; @moncrieffe1998reconceptualizing], reduces the opportunities for rent-seeking and corruption [@deininger2005does; @wenar2006accountability] and improves the quality of public services [@adsera2003you; @bjorkman2009power]. Recent research also suggests that accountability leads to higher economic growth because it limits state discretion in the economy and increases long-term investments in human capital [@benhabib2010economic; @suebvises2018social; @ponzetto2018social].

However, accountability systems take many forms. One promising model is that of _bottom-up monitoring_, in which citizens receive information about the shortcomings of a given project so they can evaluate and pressure underperforming public officials [@kosack2014does; @molina2016community; @raffler2018weakness]. Proponents argue bottom-up accountability is effective because: 1) constituents have first-hand information about the outcomes of local policies; 2) citizens have incentives to attack corruption that directly affects themselves; 3) policy-makers are sensitive to social punishment from their own communities [@serra2011combining, 570]. In this regard, bottom-up accountability offers a potential solution to the principal-agent dilemma in public service by aligning the interests of state officials with those of the constituency they serve [@barro1973control; @raffler2018weakness, 2]. [XXXX However não tá howevering... XXXX]

Here we assess the impact of _Tá de Pé_ (TDP), a mobile phone application designed to lower the costs of evaluating public works and punishing political representatives in Brazil.[XXXX] Developed by Transparência Brasil[^transparencia], TDP allows citizens to know the location of public school construction sites, check their completion status and anonymously request information from competent authorities. TDP users can also take pictures of the construction sites and submit them to independent engineers for examination. If the engineers classify the construction as delayed, TDP prompts users to send a message to the mayor's office asking for completion estimates and explanations about the construction status. TDP has been online since April 2017 in XXXX Brazilian municipalities and was the winner of the 2016 Google Social Impact grant with more than 200,000 popular votes[^google].

@@

We use the TDP app to conduct two experimental interventions and test its impact on five outcomes related to school completion rates and complaints to public authorities. Overall, our results show that providing information to citizens has no consistent impact on policy outcomes. We find that the TDP app increased the likelihood of school completion by 7 percent in the first experiment, but the effect is null in the second intervention. We also find that the app increased the percentage of the reported construction completion by 1.75 percent in our second experiment, but only after nearest neighbourhood matching adjustment. Importantly, both interventions have no significant impact on any of the other measures of school completion we employed in this study. [XXXX TOPIC Phrase: our results show that providing information to citizens has no consistent impact on policy outcomes.!!! XXXX]

The results raise questions about the ability of citizens to hold representatives accountable using bottom-up strategies. Although studies like @bjorkman2009power, @fiala2017social and @reinikka2005fighting report better policy outcomes after providing information to local communities, our findings are closer to @banerjee2010pitfalls, @keefer2014mass, @lieberman2014does, @bjorkman2017experimental, @olken2007monitoring and @raffler2018weakness offering little evidence that information-based interventions lead to greater government responsiveness. Our analysis suggests that grassroots monitoring is ineffective in altering government behaviour in Brazil, and other models of political accountability may be more effective for public service provision. [XXXX CORTAR ISSO: and other models of political accountability may be more effective for public service provision XXXX]

@@

The _Tá de Pé_[^tdp] (TDP) cell phone application is an initiative carried out by Transparência Brasil to foster bottom-up accountability in the Brazilian public sector. More specifically, the main goal of the TDP project is to improve responsiveness in government education expenditures. The TDP app incentivises citizens to provide up-to-date information about unfinished school constructions in their neighbourhoods, and that information will be assessed by a group of independent specialists. In the case the construction is behind schedule, TDP provides a writing platform whereby citizens can report to public officials quickly and anonymously. The app then prompts citizens to write a notification to the mayor's office, which has 15 days to reply. If they do not respond to the request, the app forwards the notification to the Brazilian federal government office that grants the funds for the construction, making it harder for the municipality to access federal funds in the future. The motivation behind this intervention is that bottom-up pressure from citizens empowers individuals to closely monitor public works. This, in turn, results in better social outcomes as public agents become more responsive to community demands.

@@

Transparencia Brazil built the app in the first three months of 2017 and tested it in May 2017. The first stable Android version was deployed on Google Play on 14 August 2017. A version for iOS came about six months later. In October 2017, Transparencia Brazil started a Facebook campaign in order to publicise the app. Facebook is the most widely used social media network in Brazil with around 72 million users in 2017 [@statista2019facebook]. The campaigns cost US$ 4,479.29 and Facebook advertisements attracted 2,028 new users to the platform in October 2017 only.

@@

\begin{figure}
     \begin{minipage}[t]{.3\textwidth}
       \centering
       \includegraphics[scale=0.25]{tdp03.jpg}
       \label{}
     \end{minipage}
     \hfill
     \begin{minipage}[t]{.3\textwidth}
       \centering
       \includegraphics[scale=0.25]{tdp01.jpg}
     \end{minipage}
     \hfill
     \begin{minipage}[t]{.3\textwidth}
       \centering
       \includegraphics[scale=0.25]{tdp02.jpg}
     \end{minipage}
\caption{The \textit{Tá de Pé} mobile phone application. The first image presents a list of school construction sites closer to the users' location. The second image shows that the school construction selected by the user is delayed by 9 months. The last image shows how citizens can add information to the photos they submit via the app.} 
\label{fig:screenshots}
\end{figure}

Transparência Brasil partnered with the Brazilian branch of _Engineers without Borders_ (EWB), an independent non-governmental organisation[^ewb], to provide technical assessment of school completion rates based on user-submitted photos and GPS coordinates. The engineers' reports are later uploaded to the TDP database and stored on the users' mobile devices so citizens can follow the progress of the reported constructions.

@@

The TDP has also received feedback from Brazilian computer programmers and policy analists. In 2017 and 2018, Transparência Brasil announced two team programming competitions, called 'Tá de Pé Hackathons', where contributors could fix code bugs and suggest new functionalities to the TDP project. One of these innovations consists of a Twitter bot (<https://twitter.com/tadepeapp>) which posts a message on the social network each time a user submits a new picture for evaluation or a municipality responds to a citizen's request. This allows any interested parties, including those who do not use the TDP app, to check the state of school construction sites.

@@

Between August 2017 and February 2019, we implemented two interventions to measure the effect of the TDP app on five school construction outcomes plus a placebo outcome. The outcomes are: 1) a placebo outcome indicating the percentage of the project completed before the impact evaluation started; 2) the percentage of the project completed by the end of the interventions; 3) the difference between the percentage reported as completed before and after the interventions; 4) the number of finished constructions; 5) the number of cancelled constructions; 6) the number of schools where construction companies updated the conclusion dates. In terms of expectations, Table \ref{tableOuctExp} depicts the expected effects for each of the studied outcomes.

\begin{table}[htb!]
\centering
\caption{Outcomes and expected effect of the TDP intervention}
\label{tableOuctExp}
\resizebox{\textwidth}{!}{%
\begin{tabular}{lcl}
\hline
\multicolumn{1}{c}{\textbf{}} & \textbf{\begin{tabular}[c]{@{}c@{}}Expected\\ impact\end{tabular}} & \multicolumn{1}{c}{\textbf{Meaning}} \\ \hline
\begin{tabular}[c]{@{}l@{}}Placebo:\\ Percentage of the project completed before\\ the impact evaluation started\end{tabular} & Null & \begin{tabular}[c]{@{}l@{}}The placebo outcome, as reported before the intervention, \\ should have a null impact. This represents the absence of\\ differences between treatment and control prior to the\\ intervention.\end{tabular} \\
\begin{tabular}[c]{@{}l@{}}Outcome 1:\\ Percentage of the project reported as completed\\ by the end of the intervention period\end{tabular} & Positive & \begin{tabular}[c]{@{}l@{}}If the intervention has a positive effect, the firms should\\ increase their efforts toward finishing the construction\\ more quickly.\end{tabular} \\
\begin{tabular}[c]{@{}l@{}}Outcome 2:\\ Difference between the percentage reported as\\ completed before and after the intervention\end{tabular} & Positive & \begin{tabular}[c]{@{}l@{}}If the intervention has a positive impact, the difference\\ between before and after the intervention should reflect\\ this.\end{tabular} \\
\begin{tabular}[c]{@{}l@{}}Outcome 3:\\ Dummy indicator for finished constructions\end{tabular} & Positive & \begin{tabular}[c]{@{}l@{}}If the intervention has a positive effect, more schools\\ should be reported as finished in the treatment group.\end{tabular} \\
\begin{tabular}[c]{@{}l@{}}Outcome 4:\\ Dummy indicator for cancelled construction\end{tabular} & Negative & \begin{tabular}[c]{@{}l@{}}If the intervention has a positive impact, less \\ constructions should be abandoned and reported as \\ finished in the treatment group.\end{tabular} \\
\begin{tabular}[c]{@{}l@{}}Outcome 5:\\ Number of schools where construction \\ companies updated the conclusion dates\end{tabular} & Positive & \begin{tabular}[c]{@{}l@{}}If the intervention has a positive effect, firms and \\ mayor's offices should be responsive to the public,\\ and update their finishing dates.\end{tabular} \\ \hline
\end{tabular}%
}
\end{table}

The first intervention was carried out from August 2017 to February 2018 using the Android version of TDP. The randomization was conduct at the municipal level. We randomly selected `r as.numeric(table(impEvalph1$treatOriginal)[1])` municipalities to the control group and included `r as.numeric(table(impEvalph1$treatOriginal)[2])` in the treatment group. Our control condition consists in removing all information about school construction from the TDP app in the chosen municipalities, so that citizens were unable to report constructions in the control municipalities.

@@

Table \ref{tab:results1} summarises the main results of intervention 1. Each column represents the treatment effect of the TDP app on one of the outcomes we measured for this study. All models reported here include the five control variables described in the previous section and Brazilian states' fixed effects. We also estimated the models without control variables, without fixed effects and with nearest-neighbour matching. The results are very similar to those below.

@@

We find that the intervention is sufficient to improve school completion rates. The TDP application increases in `r round(100*resInt1[[2]][[16]]$coefficients[1], 1)` percent the chance that a school construction is reported as finished by the end of the intervention. This is in line with our theoretical expectations, but as we discuss below, the finding is not consistent. All other coefficients are not statistically significant at conventional levels. On the one hand, the results indicate that our placebo outcome, the percentage of the invested executed before the intervention, indeed behaves as predicted. On the other hand, we expected the four remaining outcomes to improve after the introduction of the app. The literature on bottom-up accountability argues that more information about the shortcomings of public service provision will mobilise citizens who, in turn, would pressure state agents to produce better social policies. Our results do not lend support to that hypothesis.

We observe a similar pattern in the second intervention. Table \ref{tab:results2} shows that the treatment does not have a statistically significant effect on any our outcomes of interest, including the placebo. This provides evidence that our previous findings are not a statistical fluke, and questions the robustness of the effect of the TDP app on school completion. Nevertheless, in our additional tests we find that the app improved the difference between the baseline and the end-line reported investment by 1.57 percent. This effect only reaches statistical significance when we employ inverse probability weights with block randomisation and one-to-one nearest neighbourhood matching.

@@

We also note that the effect sizes are small in each of our original models. Therefore, even if the treatment effects were statistically significant, the TDP app would have a reduced impact over the outcomes. This has an important substantive meaning, as it poses a direct challenge to the bottom-up accountability theory, at least for the school construction outcomes we investigated here.

@@

What factors, then, are driving these results? It seems unlikely that the null results derive from flaws in the research design. First, our study is well powered. Although the treatment is indirect, meaning that the person has to download the app, find a school construction, and then report it, we included a substantial number of municipalities in treatment groups. We should be able to capture small average treatment effects if they exist. Second, balance and manipulation tests indicate that the treatment allocation was successful, so we can rule out problems in the randomisation procedures. Third, after doing a series of robustness tests --- including matching and randomised inference --- we still find no firm evidence of treatment effect. Finally, note that the sign of the coefficients are frequently contrary to our theoretical expectations. This rules out a possible concern about statistical power with our small control group approach.

@@

In sum, our experiments suggest that popular participation and bottom-up monitoring may not be effective to improve public service delivery in the case of school constructions in Brazil. Nevertheless, the null findings are informative to researchers and policy-makers. The most important recommendation derived from this study is that interventions targeting elite groups, such as lobbyists or civil servants, may render better policy outcomes than those focused at the grassroots levels. Another core lesson is that although digital interventions are promising means to deliver information, perhaps they do not have the same impact as personal, face-to-face communication. Since many developing countries share Brazil's issues with education provision, the shortcomings we describe here serve as warnings for future interventions.
